---
title: "R Workshop - MISA"
author: "Dr. J. Tichon"
date: "Feb 17, 2022"
output: 
  beamer_presentation:
    theme: "Berlin"
    colortheme: "dolphin"
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Getting Started in R

### What is R?

R is a statistical programming language that excels at tasks that involve data management, statistical analysis, data visualization,and simulation.

### What is R?

However, R is so much more than a programming language:

- Report writing
- Data management
- Creating dashboards
- Creating HTML pages
- Powerful data visualization tools
- Interactive web based demonstrations

### Who Uses R?

- Statisticians
- Any discipline that touches data 
- Front-end developers
- News sites
- E-commerce businesses
- Data analytics

### Why Should You Learn R?

- Business of data management, cleaning, and organization is quicker and more powerful
- Creates reports that seamlessly integrate text, math, code, graphs 
- Easy to share with people that require different formats: csv , html, PDF, Word
- Free, heavily supported, and found throughout every discipline and increasingly major businesses
- Creates accessible, and powerful dashboards for internal use and for websites

### Talk Outline

- Company/Business Dashboards
- Data Visualization 
- Educational Possibilities
- Getting Setup with R
- Personal Documents
- Working with Data

### Company/Business Dashboards

- [Shiny Gallery](https://shiny.rstudio.com/gallery/)
- [Career Pathfinder](https://shiny.rstudio.com/gallery/career-pathfinder.html)
- [Freedom of Press](https://shiny.rstudio.com/gallery/freedom-press-index.html)

### Data Visualization

- [Tidy Tuesday Github](https://github.com/rfordatascience/tidytuesday)
- [#TidyTuesday](https://twitter.com/search?q=%23tidytuesday&src=typed_query)
- [BBC and ggplot](https://medium.com/bbc-visual-and-data-journalism/how-the-bbc-visual-and-data-journalism-team-works-with-graphics-in-r-ed0b35693535)

### Educational Possibilities

- [Permutation Lab](https://jenna-tichon.shinyapps.io/Permutation-Tests/)
- [Regression Tutorial with Learnr](https://jenna-tichon.shinyapps.io/RegressionTutorial/)
- [Slope Simulation](https://jenna-tichon.shinyapps.io/Slope-Distribution/)

## Getting Setup with R

### Installing Software

1- Download R (https://cran.r-project.org/mirrors.html)

2- Download R Studio (https://www.rstudio.com/products/rstudio/download/)

3- Install tinytex by running the following two lines of code in console:

install.packages('tinytex')

tinytex::install_tinytex()

### R Studio Cloud

There is a web based cloud version of R Studio at (http://rstudio.cloud)

Every account has 25 free hours a month, can purchase additional time for various price points.

### Opening Up R Studio

- Take a look at the basics of R Studio
- [YouTube tutorial of main parts of the program](https://youtu.be/NgoGsL6ZjH4)

### Projects for Workflow

- For every new task you work on in R Studio, you can create a project rooted to a folder so that it always looks how you left it.

## Personal Documents

### .r files

- .r files function as plain text documents in which you can write code without having to worry about it running
- See anova.R

### Slides 

- Presentations can be made in Beamer
- These slides are simple but you can do anything you'd normally do plus more like...

### Slides - Typset math

$$\int_{0}^4\dfrac{e^{-x-2}}{4x}$$

### Slides - Include code



```{r, echo = TRUE}
cars.lm <- lm(mpg~wt, data = mtcars)
cars.lm
```

### Include Plots with Code

```{r, echo = TRUE, out.width="80%"}
plot(mpg~wt, data = mtcars,
     xlab = "Weight (tonnes)", ylab = "Mileage (mpg)")
```

### Include Plots Without Code

```{r}
plot(mpg~wt, data = mtcars,
     xlab = "Weight (tonnes)", ylab = "Mileage (mpg)")
```

### Create HTML, PDF, or Word Document

- R Markdown (.rmd) files combine math typeseting (LaTeX), r coding, and Markdown formatting to create PDF, HTML, or Word documents.
- See sports-project_J-Tichon.rmd

## Working with Data

### Writing Data Directly with Vectors

```{r, echo = TRUE}
x <- c(4, 7, 8, 9)
mean(x)
```

### Writing Multiple Vectors to a Dataframe

```{r, echo = TRUE}
name <- c("Jean", "Jorts")
colour <- c("Tortie", "Orange")
weight <- c(8, 12)
dislikes <- c("Union Busting", "Being Buttered")
catInfo <- data.frame(name, colour, weight, dislikes)
```

### Writing Multiple Vectors to a Dataframe

```{r, echo = TRUE}
catInfo
```

### Importing and Cleaning Data from a .csv File

- See census-code-clean.R

### Good Resources for Starting

- [R For Data Science](https://r4ds.had.co.nz/)
- [R For the Rest of Us](https://rfortherestofus.com/)

